---
title: "Untitled"
author: "Ji Won"
date: "4/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(maps)
library(maptools)
library(mgcv)
library(extRemes)
library(ROCR)
```


```{r, echo = FALSE }
rm(list= ls())
load("~/Desktop/Project/data_train.RData")
data_train_DF<- data_train_DF %>%
  mutate(wind = sqrt(clim1^2 + clim2^2)) %>%
  dplyr::rename(dew_temp = clim3,
         temp = clim4,
         pot_evap = clim5,
         solar_rad = clim6,
         thermal_rad = clim7,
         pressure = clim8,
         evap = clim9,
         precip = clim10) %>%
  subset(select = -c(clim1, clim2) )

dat_scaled <- as.data.frame(scale(data_train_DF[,c(26:36)]))
dat_scaled <- cbind(data_train_DF[,c(1:25)], dat_scaled)

dat_scaled$date <- with(data_train_DF, sprintf("%d-%02d", year, month))
odd_years <- filter(dat_scaled, year %%2 ==1)

smp_size <- floor(0.75 * nrow(odd_years))

set.seed(123)
train_ind <- sample(seq_len(nrow(odd_years)), size = smp_size)

train_data <- odd_years[train_ind, ]
test <- odd_years[-train_ind, ]


train_data<- train_data%>%
  group_by(lon, lat, month)%>%
  mutate(mean_monthly_BA= mean(CNT))

for (m in 3:9){
  for (L in seq(-124.75, -66.75, 0.5)){
    for (l in seq(25.25, 49.25, 0.5)){
      test$mean_monthly_BA[test$lon==L & test$lat==l & test$month==m] <- 
        train_data$mean_monthly_BA[train_data$lon==L & train_data$lat==l & train_data$month==m][1]
    }
  }
}

rm(L, l, m, data_train_DF)

```


```{r}
threshold <- 2000
```


# LOGISTIC REGRESSION FOR (BA > THRESHOLD)

## glm

```{r}

train_data$overthreshold <- ifelse(train_data$BA > threshold, 1, 0)
test$overthreshold <- ifelse(test$BA > threshold, 1, 0)

overthreshold_glm <- glm(formula = overthreshold ~ CNT + month + year + lat + thermal_rad + altiSD + solar_rad +  wind +altiMean + lc18 + lc1 + lc7 + lc8 +lc15 + lc5 + lc12 + mean_monthly_BA, family = binomial, data = train_data)

summary(overthreshold_glm)


overthresholdpredict <- predict(overthreshold_glm, newdata = test, type = "response") #predicting prb that BA > threshold
head(overthresholdpredict)

test$overthresholdpredict <- overthresholdpredict


filter(test, overthresholdpredict > 0.10)


rm(table_mat, accuracy_Test)


```

## gam

```{r}
overthreshold_bam <- bam(overthreshold ~ s(CNT) + te(lat,lon) + s(temp) + s(pressure) + s(precip) + s(pot_evap)+ s(solar_rad) + s(wind), data = train_data, method = "REML", family = binomial)

summary(overthreshold_bam)

overthresholdpredict1 <- predict(overthreshold_bam, newdata = test, type = "response") #predicting prb that BA > threshold
head(overthresholdpredict1)

test$overthresholdpredict1 <- overthresholdpredict1


table_mat1 <- table(test$overthreshold, overthresholdpredict1 > 0.10)
table_mat1

filter(test, overthresholdpredict1 > 0.14)

accuracy_Test1 <- sum(diag(table_mat1)) / sum(table_mat1)
accuracy_Test1

rm(table_mat, accuracy_Test)

```

## random forest


```{r}
library(rpart)
library(rpart.plot)
library(caTools)
library(party)
library(partykit)
library(randomForest)


r_forest <- randomForest(as.factor(overthreshold) ~ month + year + lon + lat + thermal_rad +
    area + altiSD + temp + pressure + lc1 + lc2 + lc5 + lc6 +
    lc7 + lc8 + lc9 + lc11 + lc12 + lc13 + lc15 + lc16 + lc18 , data = train_data, ntree=1000, importance=TRUE)
r_forest
rforestpredict <- predict(r_forest, test)

importance(r_forest, type=1)
varImpPlot(r_forest)

rforestprob = predict(r_forest, test, type = "prob")

test$rforestpredict <- rforestpredict
test$rforestprob <- rforestprob
test %>% filter(rforestpredict == 1)
```


# Predicting

## GPD models for predicting extremes

### vglm

```{r}
library(VGAM)

modelvglm <- vglm(BA~lon+area+temp+solar_rad+pressure+dew_temp+altiMean+lc2+lc7+lc8+lc11+lc15+lc18, family = gpd(threshold = threshold), link = "log", data = train_data, subset = BA > threshold)

summary(modelvglm)

#selected
modelvglm1 <- vglm(BA ~ solar_rad + lc2 + evap + lc15 + year, family = gpd(threshold = threshold), link = "log", data = train_data, subset = BA > threshold)

summary(modelvglm1)


```


### vgam
```{r}

```


## Predicting burnt area

### GLM
```{r}
m_glm <- glm(log(BA+1)~lat+lon+thermal_rad+ altiMean+ altiSD+temp+ solar_rad+pressure+ lc1+lc2+lc3+lc5+lc6+lc7+lc8+lc11+lc12+lc13+lc15+lc16+lc18, family= "gaussian", data = train_data)
```


### BAM
```{r}

m_bam <- bam(log(BA+1) ~ te(lat,lon) + s(temp) +s(dew_temp) + s(pressure) + s(precip) + s(pot_evap)+s(solar_rad) + s(altiMean, k = 20) + s(area) +s(lc1) + s(lc2, k = 20) + s(lc5)+s(lc7)+  s(lc9)+ s(lc11, k = 20) + s(lc12, k = 20) + s(lc13) + s(lc14) + s(lc16)+ s(lc17)+s(lc18, k = 20), data = train_data, method = "REML", family = gaussian)
```


### GBM
```{r}
m_gbm <- gbm(log(BA+1) ~ .-CNT-date-overthreshold, data = train_data, distribution = "gaussian",interaction.depth = 1, n.minobsinnode = 2,shrinkage = 0.0001)
```





# Getting predictions

## Using glm logistic & vglm GPD for over threshold & glm gaussian for BA

```{r}
new_glm <- test %>% filter(overthresholdpredict > 0.15) 
rows_glm <- which(test$overthresholdpredict > 0.15)

predictedvglm <- predict(modelvglm, untransform = TRUE, newdata = new_glm) #predict with gpd the selected rows

u_ba_threshold <- subset(u_ba, u_ba > threshold)

x = matrix(0, nrow(new_glm), length(u_ba_threshold))
for (i in 1:nrow(predictedvglm)){
  for (j in 1:length(u_ba_threshold))
    x[i,j] <- pgpd(u_ba_threshold[j], scale = predictedvglm[i,1] , shape = predictedvglm[i,2]) 
    #prob that BA <  u_ba_threshold given that BA > threshold
}

prob_threshold = matrix(0, nrow(x), length(u_ba_threshold))
for (i in 1:nrow(x)){
  prob_threshold[i,] <- x[i,] * new_glm$overthresholdpredict[i]
}

p_glm <- predict_ba_prob_gaussian_1(m_glm, test)

s <- get_score_ba(p_glm, test$BA, u_ba, weights_ba)

#0.052333 without extremes
#4186.64


p_glm[rows_glm, 22:28] <- p_glm[rows_glm, 21] +prob_threshold

p1 <- p_glm
q <- p_glm[rows_glm,28] -1
p_glm[rows_glm,] <- p1[rows_glm, ]- q

for (i in 1:24521){
  for(j in 1:28){
    if (p_glm[i,j] < 0) {
      p_glm[i,j] <- 0
    }
  }
}

score_glm <- get_score_ba(p_glm, test$BA, u_ba, weights_ba)



```



## Using bam logistic & vglm GPD for over threshold & glm gaussian for BA

```{r}
new_bam <- test %>% filter(overthresholdpredict1 > 0.15) #rows where prb exceed in threshold > 0.5
rows_bam <- which(test$overthresholdpredict1 > 0.15)

predictedvglm <- predict(modelvglm, untransform = TRUE, newdata = new_bam) #predict with gpd the selected rows

u_ba_threshold <- subset(u_ba, u_ba > threshold)

x = matrix(0, nrow(new_bam), length(u_ba_threshold))
for (i in 1:nrow(predictedvglm)){
  for (j in 1:length(u_ba_threshold))
    x[i,j] <- pgpd(u_ba_threshold[j], scale = predictedvglm[i,1] , shape = predictedvglm[i,2]) 
    #prob that BA <  u_ba_threshold given that BA > threshold
}

prob_threshold = matrix(0, nrow(x), length(u_ba_threshold))
for (i in 1:nrow(x)){
  prob_threshold[i,] <- x[i,] * new_bam$overthresholdpredict1[i]
}

p_bam <- predict_ba_prob_gaussian_1(m_glm, test)

p_bam[rows_bam, 22:28] <- p_bam[rows_bam, 21] +prob_threshold

p1 <- p_bam
q <- p_bam[rows_bam,28] -1
p_bam[rows_bam,] <- p1[rows_bam, ]- q

for (i in rows_bam){
  for(j in 1:28){
    if (p_bam[i,j] < 0) {
      p_bam[i,j] <- 0
    }
  }
}
score_bam <- get_score_ba(p_bam, test$BA, u_ba, weights_ba)


```


## Using random forest & vglm GPD for over threshold & glm gaussian for BA

```{r}
new_rforest <- test %>% filter(rforestpredict == 1) #rows where prb exceed in threshold > 0.5
rows_rforest <- which(rforestpredict == 1)

predictedvglm <- predict(modelvglm, untransform = TRUE, newdata = new_rforest) #predict with gpd the selected rows

x = matrix(0, nrow(new_rforest), length(u_ba_threshold))
for (i in 1:nrow(predictedvglm)){
  for (j in 1:length(u_ba_threshold))
    x[i,j] <- pgpd(u_ba_threshold[j], scale = predictedvglm[i,1] , shape = predictedvglm[i,2]) 
    #prob that BA <  u_ba_threshold given that BA > threshold
}

prob_threshold = matrix(0, nrow(x), length(u_ba_threshold))
for (i in 1:nrow(x)){
  prob_threshold[i,] <- x[i,] * new_rforest$rforestprob[i]
}

p_rforest <- predict_ba_prob_gaussian_1(m_glm, test)


p_rforest[rows_rforest, 22:28] <- p_rforest[rows_rforest, 21] + prob_threshold

p1 <- p_rforest
q <- p_bam[rows_bam,28] -1
p_rforest[rows_rforest,] <- p1[rows_rforest, ]- q

for (i in rows_rforest){
  for(j in 1:28){
    if (p_rforest[i,j] < 0) {
      p_rforest[i,j] <- 0
    }
  }
}
score_rforest <- get_score_ba(p_rforest, test$BA, u_ba, weights_ba)
#0.05285 with extremes
```



## Using bam logistic & vglm GPD for over threshold & bam gaussian for BA (score_bam1)
```{r}
new_bam <- test %>% filter(overthresholdpredict1 > 0.15) #rows where prb exceed in threshold > 0.5
rows_bam <- which(test$overthresholdpredict1 > 0.15)

predictedvglm <- predict(modelvglm, untransform = TRUE, newdata = new_bam) #predict with gpd the selected rows

u_ba_threshold <- subset(u_ba, u_ba > threshold)

x = matrix(0, nrow(new_bam), length(u_ba_threshold))
for (i in 1:nrow(predictedvglm)){
  for (j in 1:length(u_ba_threshold))
    x[i,j] <- pgpd(u_ba_threshold[j], scale = predictedvglm[i,1] , shape = predictedvglm[i,2]) 
    #prob that BA <  u_ba_threshold given that BA > threshold
}

prob_threshold = matrix(0, nrow(x), length(u_ba_threshold))
for (i in 1:nrow(x)){
  prob_threshold[i,] <- x[i,] * new_bam$overthresholdpredict1[i]
}

p_bam1 <- predict_ba_prob_gaussian_1(m_bam, test)
s1 <- get_score_ba(p_bam1, test$BA, u_ba, weights_ba)

p_bam1[rows_bam, 22:28] <- p_bam1[rows_bam, 21] +prob_threshold

p1 <- p_bam1
q <- p_bam1[rows_bam,28] -1
p_bam1[rows_bam,] <- p1[rows_bam, ]- q

for (i in rows_bam){
  for(j in 1:28){
    if (p_bam1[i,j] < 0) {
      p_bam1[i,j] <- 0
    }
  }
}
score_bam1 <- get_score_ba(p_bam1, test$BA, u_ba, weights_ba)

```


## Using bam logistic & vglm GPD for over threshold & gbm gaussian for BA (score_bam2)
```{r}
new_bam <- test %>% filter(overthresholdpredict1 > 0.15) #rows where prb exceed in threshold > 0.5
rows_bam <- which(test$overthresholdpredict1 > 0.15)

predictedvglm <- predict(modelvglm, untransform = TRUE, newdata = new_bam) #predict with gpd the selected rows

u_ba_threshold <- subset(u_ba, u_ba > threshold)

x = matrix(0, nrow(new_bam), length(u_ba_threshold))
for (i in 1:nrow(predictedvglm)){
  for (j in 1:length(u_ba_threshold))
    x[i,j] <- pgpd(u_ba_threshold[j], scale = predictedvglm[i,1] , shape = predictedvglm[i,2]) 
    #prob that BA <  u_ba_threshold given that BA > threshold
}

prob_threshold = matrix(0, nrow(x), length(u_ba_threshold))
for (i in 1:nrow(x)){
  prob_threshold[i,] <- x[i,] * new_bam$overthresholdpredict1[i]
}

p_bam2 <- predict_ba_prob_gaussian_2(m_gbm, test, "response" )
s2 <- get_score_ba(p_bam2, test$BA, u_ba, weights_ba)

p_bam2[rows_bam, 22:28] <- p_bam2[rows_bam, 21] +prob_threshold

p1 <- p_bam2
q <- p_bam2[rows_bam,28] -1
p_bam2[rows_bam,] <- p1[rows_bam, ]- q

for (i in rows_bam){
  for(j in 1:28){
    if (p_bam2[i,j] < 0) {
      p_bam2[i,j] <- 0
    }
  }
}
score_bam2 <- get_score_ba(p_bam2, test$BA, u_ba, weights_ba)

```