---
title: "GAMs"
author: "Ji Won"
date: "3/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mgcv)
library(dplyr)
library(tidyverse)
library(caret)
```


```{r, echo=FALSE}
rm(list = ls())
load("~/Desktop/Project/data_train.RData")
data_train_DF<- data_train_DF %>%
  mutate(wind = sqrt(clim1^2 + clim2^2)) %>%
  dplyr::rename(dew_temp = clim3,
         temp = clim4,
         pot_evap = clim5,
         solar_rad = clim6,
         thermal_rad = clim7,
         pressure = clim8,
         evap = clim9,
         precip = clim10
  ) %>%
  subset(select = -c(clim1, clim2) )



dat_scaled <- as.data.frame(scale(data_train_DF[,c(26:36)]))
dat_scaled <- cbind(data_train_DF[,c(1:25)], dat_scaled)

odd_years <- filter(dat_scaled, year %% 2 == 1)
dat_scaled$fire <- ifelse(dat_scaled$CNT>0, 1, 0)

train_data <- filter(dat_scaled, year == c(1995, 2005, 2015, 2001, 1999))
odd_years <- filter(dat_scaled, year %%2 ==1)
test <- filter(dat_scaled, year == c(1997,2007, 2003, 2011))
```


## Logistic BAMs for fire classification


```{r pressure, echo=FALSE}
library(car)
m <- bam(fire ~ s(lat,lon, k = 50)+s(month, k=7) + s(temp) + s(pressure) + s(precip) + s(wind)+s(pot_evap)+s(solar_rad) + s(altiSD) + s(area) +s(lc1) + s(lc2) + s(lc3) + s(lc4) + s(lc5)+s(lc7)+ s(lc8)+ s(lc9)+ s(lc11, k = 20) + s(lc12, k = 20) + s(lc13) + s(lc14) + s(lc15) + s(lc16, k = 6)+ s(lc17)+s(lc18), data = train_data, method = "REML", family = binomial)
summary(m)
gam.check(m)
plot(m, shade = TRUE, rug = TRUE)
```


```{r}
m2 <- bam(fire ~ s(lat,lon, k = 100) + s(temp) +s(dew_temp) + s(pressure) + s(precip) + s(pot_evap)+s(solar_rad) + s(altiMean) + s(area) +s(lc1) + s(lc2) + s(lc5)+s(lc7)+  s(lc9)+ s(lc11, k = 20) + s(lc12, k = 20) + s(lc13) + s(lc14) + s(lc16)+ s(lc17)+s(lc18), data = train_data, method = "REML", family = binomial)
summary(m2)
gam.check(m2)
plot(m2, shade = TRUE, rug = TRUE)
```




```{r}
m3 <- bam(fire ~ te(lat,lon) + s(temp) +s(dew_temp) + s(pressure) + s(precip) + s(pot_evap)+s(solar_rad) + s(altiMean, k = 20) + s(area) +s(lc1) + s(lc2, k = 20) + s(lc5)+s(lc7)+  s(lc9)+ s(lc11, k = 20) + s(lc12, k = 20) + s(lc13) + s(lc14) + s(lc16)+ s(lc17)+s(lc18, k = 20), data = train_data, method = "REML", family = binomial)
summary(m3)
gam.check(m3)

```

```{r}
m4 <- bam(fire ~ te(lat,lon) + s(temp) +s(dew_temp) + s(pressure) + s(precip) + s(pot_evap)+s(solar_rad)+ s(area) +s(lc1) + s(lc2, k = 20) + s(lc5)+s(lc7)+  s(lc9)+ s(lc11, k = 20) + s(lc12, k = 20) + s(lc13) + s(lc14) + s(lc16, k = 22)+ s(lc17)+s(lc18, k = 30), data = train_data, method = "REML", familiy = binomial)
summary(m4)
gam.check(m4)

```

```{r}
AIC(m)
AIC(m2)
AIC(m3)
AIC(m4)
```

```{r}

test_data <- dat_scaled %>% filter(year == c(1997,2007))

predicted2007 <- predict(m3,newdata = data_predict,type="response")
head(predicted2007)


pred = rep(0, dim(data_predict)[1])
pred[predicted2007 > .5] = 1

table(pred)
table(data_predict$fire)

```



```{r pressure, echo=FALSE}
library(car)
m <- bam(CNT ~ s(lat,lon, k = 50)+s(month, k=7) + s(temp) + s(pressure) + s(precip) + s(wind)+s(pot_evap)+s(solar_rad) + s(altiSD) + s(area) +s(lc1) + s(lc2) + lc3 + lc4 + s(lc5)+s(lc7)+ s(lc8)+ s(lc9)+ s(lc11, k = 20) + s(lc12, k = 20) + s(lc13) + lc14 + lc15 + lc16+lc18, data = train_data, method = "REML", family = poisson)
summary(m)
gam.check(m)
plot(m, shade = TRUE, rug = TRUE)
```


```{r}
library(gbm)
gbm1 <- gbm(CNT ~.-BA, data = train_data, distribution = "poisson", n.trees = 500, shrinkage = 0.1,             
            interaction.depth = 3, bag.fraction = 0.5, train.fraction = 0.5,  
            n.minobsinnode = 10, cv.folds = 5, keep.data = TRUE)  
best.iter <- gbm.perf(gbm1, method = "cv")
print(best.iter)
p_gbm1 <- predict(gbm1, newdata = test, n.trees = best.iter, type = "response")

p_gbm1<- predict_cnt_prob_poisson(gbm1, test)

score_gbm1 <- get_score_cnt(p, test$CNT, u_cnt, weights_cnt)
```

```{r}
predict_ba_prob_gaussian_2 = function(model, test, type){
  prediction = predict(model, test, type=type)
  ba_probabilities = matrix(0, nrow(test), length(u_ba))
  for (i in 1:nrow(test)){
    mu = prediction[i]   #$fit[i]
   #sd = prediction$se.fit[i]
    sd = sqrt(sum((mu-mean(prediction))^2)/(length(test)-1))
    ba_probabilities[i,] = pnorm(log(u_ba+1), mean = mu, sd = sd)
  }
  ba_probabilities
}
```


#GBM for ba
```{r}
ba_gbm <- gbm(log(BA+1)~.-CNT, data = train_data, distribution = "gaussian", n.trees = 500, shrinkage = 0.1,             
            interaction.depth = 3, bag.fraction = 0.5, train.fraction = 0.5,  
            n.minobsinnode = 10, cv.folds = 5, keep.data = TRUE)  

#ba_gbm_p <- predict(ba_gbm, newdata = test, n.trees = best.iter, type = "response")

p <- predict_ba_prob_gaussian_2(ba_gbm, test,"response")

score_ba_gbm <- get_score_ba(p, test$BA, u_ba, weights_ba)
```
```




#BAM for burnt area

##model1

```{r}
m <- bam(log(BA+1) ~ te(lat,lon) + s(temp) +s(dew_temp) + s(pressure) + s(precip) + s(pot_evap)+s(solar_rad) + s(altiMean, k = 20) + s(area) +s(lc1) + s(lc2, k = 20) + s(lc5)+s(lc7)+  s(lc9)+ s(lc11, k = 20) + s(lc12, k = 20) + s(lc13) + s(lc14) + s(lc16)+ s(lc17)+s(lc18, k = 20), data = train_data, method = "REML", family = gaussian)
```


```{r}
summary(m)
gam.check(m)
```

```{r}
p <- predict_ba_prob_gaussian_1(m, test)
```


```{r}
score<-get_score_ba(p, test$BA, u_ba, weights_ba)
```


##model2
```{r}
m2 <- bam(log(BA+1) ~ s(lat) + s(temp) +s(dew_temp) + s(pressure) + s(precip) + s(pot_evap)+s(solar_rad) + s(altiMean, k = 20) + s(area) +s(lc2, k = 20) + s(lc5)+s(lc7)+  s(lc9)+ s(lc11, k = 20) + s(lc12, k = 20) + s(lc13) + s(lc14) + s(lc16, k = 20)+ s(lc17)+s(lc18, k = 20), data = train_data, method = "REML", family = gaussian)
```


```{r}
p2 <- predict_ba_prob_gaussian_1(m2, test)
```


```{r}
score2<-get_score_ba(p2, test$BA, u_ba, weights_ba)
```
